---
title: "Data from WRDS in R"
author: "Bas Machielsen"
date: "3/26/2020"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)
```

## Introduction

In this document, I describe how to run queries to the WRDS database in R. In order to do that, we follow the guide from [the official WRDS website](https://wrds-www.wharton.upenn.edu/pages/support/programming-wrds/programming-r/r-from-your-computer/). Let us try to obtain the annual report data for Norways OBX, France's CAC40 and the Netherlands' AEX indices. 

## Installing the R SQL Client

We need `RPostgres` to interact with the WRDS database. User and password should be stored in a separate file in your home directory, see the guide for details. 

```{r}
#install.packages('RPostgres')
library(RPostgres)
```


## RPostgers

After creating .pgpass, to store your password and username for WRDS, we can try to connect to the database.

```{r}

wrds <- dbConnect(Postgres(),
                  host='wrds-pgdata.wharton.upenn.edu',
                  port=9737,
                  dbname='wrds',
                  sslmode='require',
                  user='machiels')

```

You will get an object called WRDS, a Large PqConnection, as a confirmation that the connection was made succesfully.

## Queries 

Now, we can start doing queries. Let's get all companies from the AEX, CAC40 and OBX (Norway), and let's try to retrieve the annual data. 

> WRDS data is organized by vendor (such as crsp) and referred to as a library. Each library contains a number of database tables or datasets (such as dsf), which contain the actual data in tabular format with column headers called variables.


First, let us look at how to browse. Which databases are there? 

```{r}
#"Main menu"
res <- dbSendQuery(wrds, "select distinct table_schema
                   from information_schema.tables
                   where table_type ='VIEW'
                   or table_type ='FOREIGN TABLE'
                   order by table_schema")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data

```

```{r}
res <- dbSendQuery(wrds, "select distinct table_name
                   from information_schema.columns
                   where table_schema='compg'
                   order by table_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data
```

```{r}
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='compg'
                   and table_name='g_idxcst_his'
                   order by column_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data

#So these are the index constituents at that point in time 
```

Now, let us select only those observations that belong to the indices that interest us: 

```{r}
a <- "select * from compg.g_idxcst_his WHERE gvkeyx = '150262' or gvkeyx = '150093' or gvkeyx = '150267' "
#Use a dot to go from database to dataset)

res <- dbSendQuery(wrds, a) 
data <- dbFetch(res, n = -1)
dbClearResult(res)
stocks <- data

```


Now, let us get back to the "main menu", and try to make a query to the compustat fundamentals annual database and use these firms. 

```{r}
#"Main menu"
res <- dbSendQuery(wrds, "select distinct table_schema
                   from information_schema.tables
                   where table_type ='VIEW'
                   or table_type ='FOREIGN TABLE'
                   order by table_schema")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data


#Let's see what's in comp
res <- dbSendQuery(wrds, "select distinct table_name
                   from information_schema.columns
                   where table_schema='comp'
                   order by table_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data

#Let's see what variables are in funda
res <- dbSendQuery(wrds, "select column_name
                   from information_schema.columns
                   where table_schema='comp'
                   and table_name='funda'
                   order by column_name")
data <- dbFetch(res, n=-1)
dbClearResult(res)
data

#Now, we can combine variables from different datasets using an extensive query
"select * from compg.g_idxcst_his WHERE gvkeyx = '150262' or gvkeyx = '150093' or gvkeyx = '150267' "

res <- dbSendQuery(wrds, "select a.*, b.conm, b.cusip, b.datadate from compg.g_idxcst_his a join comp.funda b on a.gvkey = b.gvkey where a.gvkeyx = '150262' or a.gvkeyx = '150093' or a.gvkeyx = '150267' ")
data <- dbFetch(res, n=-1)
dbClearResult(res)
stocks <- data

#Now, let's make another query to retrieve a lot of financial data. For example:
#act - current assets
#at - total assets
#dltt - long term debt
#emp - employees
#lt - total liabilities
#gdwl - goodwill
res <- dbSendQuery(wrds, "select gvkey, datadate, act, at, dltt, emp, lt, gdwl from comp.funda")
data <- dbFetch(res, n = -1)
dbClearResult(res)

#Let's now try to match the stocks that are of interest to us to the financial data
data <- data %>%
  mutate(datadate = str_extract(datadate,"[0-9]{4}"))

stocks <- stocks %>%
  mutate(datadate = str_extract(datadate, "[0-9]{4}"))

ourdata <- left_join(stocks, data, by = c("gvkey", "datadate"))


```

And here is a fully reproducible data collection strategy. We can also easily adjust the variables we want by looking up their codes and adding them to the SQL query that we made. 

Finally, we close the connection by the following command:

```{r}
dbDisconnect(conn =wrds)
```
























